---
title: "article-scripts"
author: "Rick Dean"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This document provides the R script for reproducing the article [The Trump Vote Request and Sentiment Analysis](https://github.com/deandevl/R_Trump_Vote_Request).

Second level sections in this document follow the article's second level sections. Subsections in this document describe the R script in producing the article.

```{r}
library(knitr)
library(kableExtra)
library(data.table)
library(readr)
library(stringi)
library(RtextminerPkg)
library(RplotterPkg)
library(RtsaPkg)
library(sentimentr)
library(stopwords)
library(ggplot2)
library(here)
```

## Read the data

### Read the file containing the transcript
```{r}
current_dir <- here()
data_file_path <- file.path(current_dir, "data", "trump_vote_request.txt")
linesInFile <- base::readLines(data_file_path)
```

### Create a data.table with the file statistics
```{r}
file_stats_dt <- data.table(
  Statistic = c("FileSize","FileLines","FileWords"),
  Value = c(format(object.size(linesInFile), units = "Kb"),
            length(linesInFile), sum(stringi::stri_count_words(linesInFile)))
)
```

### Create/save Table 1 "File Statistics" image
```{r}
RplotterPkg::create_table(
  x = file_stats_dt,
  caption = "File Statistics",
  font_size = 14,
  align_v = c("l","c"),
  col_spec_df = data.frame(
    column = c(1,2),
    width = c("4cm", "4cm")
  ),
  output_file = file.path(current_dir, "img","table-1-file-stats.png")
)
```

## Split transcript into sentences

### Split transcript by sentences
```{r}
sentences_lst <- RtextminerPkg::tokenize_text(
  x = linesInFile,
  type = "sentence"
)
```

### Create sentences_dt dataframe/add an index column
```{r}
n_rows <- nrow(sentences_lst$tokens_dt)
sentences_dt <- sentences_lst$tokens_dt
sentences_dt[,sentenceIdx := 1:n_rows]
```

### Create/save Table 2 "Sentences of the transcript" image
```{r}
RplotterPkg::create_table(
  x = sentences_dt[1:3,],
  caption = "Sentences of the transcript",
  font_size = 14,
  align_v = c("l","c"),
  col_spec_df = data.frame(
    column = c(1,2),
    width = c("12cm", "4cm")
  ),
  output_file = file.path(current_dir, "img","table-2-sentences.png")
)
```

## Split transcript into words and phrases

### Split the transcript by words/remove stopwords
```{r}
stop_words <- stopwords::stopwords(source = "snowball")
words_lst <- RtextminerPkg::tokenize_text(
  x = sentences_dt,
  input_col = "sentence",
  stopwords = stop_words,
  strip_numeric = T,
  strip_punct = T,
  strip_non_alphanum = T
)
```

#### Create/save Table 3, the word table image
```{r}
RplotterPkg::create_table(
  x = words_lst$tokens_dt[1:8,],
  caption = "Words",
  font_size = 14,
  align_v = c("l","c"),
  col_spec_df = data.frame(
    column = c(1,2),
    width = c("4cm","6cm")
  ),
  output_file = file.path(current_dir, "img","table-3-words.png")
)
```

### Split the transcript by two word phrases
```{r}
phrases_lst <- RtextminerPkg::tokenize_text(
  x = sentences_dt,
  input_col = "sentence",
  type = "ngram"
)
```

#### Change column name "ngram" to "phrase" in tokens data frame
```{r}
data.table::setnames(phrases_lst$tokens_dt, old = "ngram", new = "phrase")
```

#### Add a "phrase" column in the count data frame
```{r}
phrases_lst$tokens_count[, phrase := paste(token_1, token_2)]
```

#### Create/save Table 4, the phrase table image
```{r}
RplotterPkg::create_table(
  x = phrases_lst$tokens_dt[1:8,.(sentenceIdx,phrase)],
  caption = "Phrases",
  font_size = 14,
  align_v = c("l","c"),
  col_spec_df = data.frame(
    column = c(1,2),
    width = c("4cm","6cm")
  ),
  output_file = file.path(current_dir, "img","table-4-phrases.png")
)
```

## Word and phrase frequencies

### Create/save Figure 1, bar chart for words
```{r}
RplotterPkg::create_bar_plot(
  df = words_lst$tokens_count[1:20],
  aes_x = "word",
  aes_y = "N",
  title = "Word Frequencies",
  x_title = NULL,
  y_title = "Frequency",
  y_limits = c(0,160),
  y_major_breaks = seq(from = 0, to = 160, by = 20),
  rot_y_tic_label = T,
  bar_fill = "blue",
  bar_color = "gold",
  bar_alpha = 0.7,
  do_coord_flip = T,
  order_bars = "asc",
  bar_labels = T,
  png_file_path = file.path(current_dir, "img", "figure-1-words-bar.png")
)
```

### Create/save Figure 2, bar chart for phrases
```{r}
RplotterPkg::create_bar_plot(
  df = phrases_lst$tokens_count[1:22],
  aes_x = "phrase",
  aes_y = "N",
  title = "Phrase Frequencies",
  x_title = NULL,
  y_title = "Frequency",
  y_limits = c(0,160),
  y_major_breaks = seq(from = 0, to = 160, by = 20),
  rot_y_tic_label = T,
  bar_fill = "blue",
  bar_color = "gold",
  bar_alpha = 0.7,
  do_coord_flip = T,
  order_bars = "asc",
  bar_labels = T,
  png_file_path = file.path(current_dir, "img", "figure-2-phrases-bar.png")
)
```

## Interesting phrases

### Locate sentences that have "you know" and "we won" together
```{r}
we_won_dt <- sentences_dt[data.table::like(sentences_dt$sentence, "we won")]
you_know_we_won_dt <- we_won_dt[data.table::like(we_won_dt$sentence, "you know")]
```

### Create/save Table 5, "you know" + "we won" table
```{r}
RplotterPkg::create_table(
  x = you_know_we_won_dt,
  caption = "'you know' and 'we won' in Sentences",
  font_size = 14,
  align_v = c("l","c"),
  col_spec_df = data.frame(
    column = c(1,2),
    width = c("14cm", "4cm")
  ),
  output_file = file.path(current_dir, "img","table-5-you-know-we-won.png")
)
```

## Sentiment analysis of sentences

### Compute the sentiment measures
```{r}
sentence_sentiment_dt <- sentimentr::sentiment(sentences_dt$sentence)
```

### Rename a column and remove a column
```{r}
data.table::setnames(sentence_sentiment_dt, old = "element_id", new = "sentenceIdx")
sentence_sentiment_dt[, sentence_id := NULL]
```

### Create/save Figure 3, sentiment measures scatter plot
```{r}
RplotterPkg::create_scatter_plot(
  df = sentence_sentiment_dt,
  aes_x = "sentenceIdx",
  aes_y = "sentiment",
  connect = T,
  show_pts = F,
  x_limits = c(0,800),
  x_major_breaks = seq(0,800,100),
  y_limits = c(-1,1.25),
  y_major_breaks = seq(-1,1.25,0.25),
  rot_y_tic_label = T,
  title = "Sentence Measurements" ,
  subtitle = "777 sentences -- negative to positive scale",
  x_title = "Sentence Index",
  y_title = "Sentiment Scale",
  png_file_path = file.path(current_dir, "img", "figure-3-sentiment-scatter.png"),
  png_width_height = c(1060,480)
)
```

### Create/save Table 6, prominent sentences
```{r}
prominent_sent_dt <- data.table(
  Sentence = c("it was much more criminal than that.",
               "they’re really wrong, and they’re really wrong, brad.",
               "go ahead, please.",
               "and the truth, the real truth is i won by 400,000 votes, at least."),
  Score = c(-0.74, -0.85, 1.039, 1.026),
  SentenceID = c(477, 629, 400, 764)
)
RplotterPkg::create_table(
  x = prominent_sent_dt,
  caption = "Prominent Negative/Positive Sentences",
  font_size = 14,
  align_v = c("l","c","c"),
  col_spec_df = data.frame(
    column = c(1,2,3),
    width = c("14cm", "4cm", "4cm")
  ),
  output_file = file.path(current_dir, "img","table-6-prominent-sentences.png")
)
```

### Create/save Figure 4, moving average of sentiment scores
```{r}
RtsaPkg::graph_ma(
  df = sentence_sentiment_dt,
  time_col = "sentenceIdx",
  value_col = "sentiment",
  window_n = 5,
  #ma_type = "wma",
  show_observe = T,
  title = "Simple Moving Average Sentiment of Sentences",
  x_limits = c(0,800),
  x_major_breaks = seq(0,800,100),
  show_pts = F,
  col_width = 12,
  row_height = 7,
  palette_colors = c("black", "red"),
  png_file_path = file.path(current_dir, "img","figure-4-simple-ma.png")
)
```

